A saturation model for impaired learning with enhanced plasticity

### 2013-07-12 MHCslides/MHCslides.tex
expt results section

\usepackage{array}
\input{sl_slide_preamble.tex}
%\input{sl_slide_preamble_nonotes.tex}
\input{sl_slide_graphics_preamble.tex}
\input{sl_definitions.tex}
\input{sl_slide_symbols.tex}
%
\graphicspath{{"Figs/"}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\I}{\mathbf{I}}
%prob vector
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\wv}{\vec{w}}
%super/subscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
%quantities
\newcommand{\initial}{\mathcal{I}}
\newcommand{\area}{\mathcal{A}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\comp}{^\mathrm{c}}
\renewcommand{\e}{\mathsf{e}}
%---------Title-----------------------------------------------------------

@@

\begin{frame}
%
 \titlepage
%
\end{frame}

@@

 \vp Needs to be adjusted as eye muscles age, \etc
 }
%
\end{frame}

@@

\begin{frame}{VOR training}
%
 \parbox[t]{0.2\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VORinc.svg}}\\[1cm]
   \note[item]{trick brain into thinking VOR gain needs adjusting my moving visual stimulus}
   \note[item]{anti-phase \lto increase gain}
 \aligntop{\includegraphics[width=0.99\linewidth]{VORdec.svg}}}
 \note[item]{in phase \lto decrease gain}
 \hspace{0.15\linewidth}
 \parbox[t]{0.59\linewidth}{\aligntop{\includegraphics[width=0.99\linewidth]{VORcircuit.svg}}}
 \note[item]{Gain change involves cerebellum}
 \note[item]{Marr-Albus-Ito: Pf-Pk synapses}
 \note[item]{Lisberger-Miles: Vestibular input-VN synapses}
 \note[item]{Different mechs for different freq, head angle, gain up/down.}
 \note[item]{Different Pk cells have different tunings.}
 \note[item]{Gain up in case of interest: LTD in Pf-Pk in flucculus}
 \note[item]{Gain down: uses different mech for behaviour, but does reverse LTD in Pf-Pk in flucculus}
%
\end{frame}

@@

@@

\begin{frame}{Simplifying assumptions}
%
\begin{itemize}
  \item No spatial/temporal correlations in plasticity events.
      %
      \note[item]{allows us to concentrate on synapse, not neuron/network}
      %
%  \item Which synapses eligible for plasticity chosen randomly.
      %
      \note[item]{No filing system}
      %
  \item Potentiating/depressing plasticity events $\sim$ Poisson processes.% with rates $rf\potdep$, where $f\pot+f\dep=1$.
      %
      \note[item]{don't care if STDP...}
%      %
%      \note[item]{$r=$ total rate of plasticity events per synapse, $f\potdep=$ fraction of events that are potentiating/depressing.}
      %
  \item Potentiation and depression are described by Markov processes.% with transition probabilities $\M\potdep$.
%      %
%      \note[item]{matrix elements: transition prob from $i\to j$, given pot/dep}
%  \item Ideal observer: read weights directly.
%      \note[item]{not electrical activity: don't model neurons/network}
%      \note[item]{upper bound on electrical activity readout}
%      %
%  \item Synaptic weights can only take values $\pm1$.
      %
  \citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
      %
      \note[item]{looks like binary synapse from outside. Inside...}
\end{itemize}
%
\end{frame}

@@

\begin{frame}{Dynamics}
%
  There are $N$ identical synapses with $M$ internal functional states.
  %
  \begin{center}
    \includegraphics[width=0.7\linewidth]{model.svg}
  \end{center}
  %
  \note[item]{stoch process has steady state.}
  \note[item]{Prior activity puts it in this state. row vec.}
  \note[item]{plasticity events at rate r}
  \note[item]{fraction pot/dep}
  \note[item]{probs changed by Markov matrices, prob $i\to j$}
  \note[item]{Readout: synaptic weight vec when in each state.}
  \vp
% At $t=0$, the memory is created by $\M\potdep$ with probability $f\potdep$.
% \note[item]{for this one, we keep track of pot/dep, look for inc/dec of $\w$}
%
% \vp Forgetting caused by subsequent memories, evolving as
%
  \begin{equation*}
  \begin{aligned}
    \diff{\pr(t)}{t} &= r\pr(t)\frg,
    &\qquad
    \frg &= f\pot\M\pot+f\dep\M\dep-\I,\\&&
    \eq\frg &=0.
  \end{aligned}
  \end{equation*}
%
  \note[item]{Memory at $t=0$, keep track of pot/dep}
  \note[item]{subsequent: average over pot/dep}
% \note[item]{$\frg$ is forgetting matrix, $\I=$identity, don't keep track of pot/dep}
% Eventually, this will settle into the equilibrium distribution:
% %
% %
% \note[item]{In equilibrium prior to memory creation}
%
\end{frame}

@@

\begin{frame}{Modelling VOR learning}
%
\begin{tabular}{>{\bfseries}ll>{\impl}l}
    Mutation: &
    Changes mechanism of LTD  &
    change $\M\dep$.
    \note[item]{lower threshold \lto increase off-diagonal elements.}
 \\[1cm]
    Training: &
    Changes statistics of LTP/LTD  &
    change $r,f\pot,f\dep$.
    \note[item]{Only parameters we have. Don't care about $r$.}
 \\[1cm]
    Learning: &
    Change in VOR gain &
    decrease in $\av{\w}$.
    \note[item]{Only output we have. Don't keep track of synaptic identity.}
\end{tabular}
%
\end{frame}

@@

\begin{frame}{Serial synapse}
%
 %
 \begin{center}
   \includegraphics[width=0.5\linewidth]{serial.svg} \\[1cm]
 \note[item]{Still looks binary from outside. Hidden states (not essential).}
 \note[item]{Only see $\Delta w$ at boundary.}
   \aligntop{\includegraphics[width=0.35\linewidth]{serial_learnS.eps}}
   \hspace{2cm}
   \aligntop{\includegraphics[width=0.35\linewidth]{comparisons_serial.svg}}
 \end{center}
 %

@@

>>>>

\usepackage{array}
\usepackage{cancel}
\input{sl_slide_preamble.tex}
%\usecolortheme[named=darkblue]{structure}
\usecolortheme[rgb={0.6372549 0.27843137 0.27058824}]{structure}
%\setbeamercolor{alerted text}{fg=blue!80!black}
%\input{sl_slide_preamble_nonotes.tex}
\input{sl_slide_graphics_preamble.tex}
\input{sl_definitions.tex}
\input{sl_slide_symbols.tex}
%
\graphicspath{{"Figs/"}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\I}{\mathbf{I}}
%prob vector
\newcommand{\pr}{\mathbf{p}}
%equilibrium distribution
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\wv}{\vec{w}}
%super/subscripts
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
%quantities
\newcommand{\initial}{\mathcal{I}}
\newcommand{\area}{\mathcal{A}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\comp}{^\mathrm{c}}
\renewcommand{\e}{\mathsf{e}}
%---------Title-----------------------------------------------------------

@@

\begin{frame}
%
 \titlepage
 \note[item]{Acknowledge Barbara and Grace}
%
\end{frame}

\begin{frame}{Introduction}
%
 One might expect enhanced plasticity to enhance learning.
 \note[item]{It does help in some cases}
 But often it actually causes impairment.

 \vp We argue that this impaired learning is due to basal activity biasing the synapses prior to learning, leaving fewer synapses available for further plasticity.

 \vp By analysing several models of complex synapses, we see that the behaviour of enhanced LTD mice in a motor learning task can constrain the synaptic structure.
%
\end{frame}

@@

 \vp Needs to be adjusted as eye muscles age, \etc
 }

 \citerr{duLac1995review,Boyden2004review}
%
\end{frame}

@@

\begin{frame}{VOR training}
%
 \parbox[t]{0.2\linewidth}{%
   \aligntop{\includegraphics[width=0.99\linewidth]{VORinc.svg}}\\[1cm]
     \note[item]{trick brain into thinking VOR gain needs adjusting my moving visual stimulus}
     \note[item]{anti-phase \lto increase gain}
   \aligntop{\includegraphics[width=0.99\linewidth]{VORdec.svg}}
     \note[item]{in phase \lto decrease gain}
   }
 \hspace{0.15\linewidth}
 \parbox[t]{0.59\linewidth}{%
   \aligntop{\includegraphics[width=0.99\linewidth]{VORcircuit.svg}}

   \vp PF-Pk: \parbox[t]{0.5\linewidth}{PF+CF \lto LTD, \\
              PF+\cancel{CF} \lto LTP.}
 }
   \note[item]{Gain change involves cerebellum}
   \note[item]{Marr-Albus-Ito: Pf-Pk synapses}
   \note[item]{Lisberger-Miles: Vestibular input-VN synapses}
   \note[item]{Different mechs for different freq, head angle, gain up/down.}
   \note[item]{Different Pk cells have different tunings.}
   \note[item]{Gain up in case of interest: LTD in Pf-Pk in flocculus}
   \note[item]{Gain down: uses different mech for behaviour, but does reverse LTD in Pf-Pk in flocculus}

   \citerr{Marr1969cerebellum,Albus1971cerebellum,ito1972cerebellum,Ito1982VOR,Miles1981VOR}
%
\end{frame}

@@

\begin{frame}{MHC-I D$^\mathrm{b}$K$^\mathrm{b} -/-$ knockout}
%
 Knockout of molecule lowers threshold for LTD in PF-Pk synapses.\\
 \citerr{McConnell2009MHCcerebellum}
 \note[item]{Easier LTD \lto expect better learning}

 \vp \begin{center}
 \begin{tabular}{ccc}
   Baseline & \hspace{0.1\linewidth} & Gain increase at $1.0\,\mathrm{Hz}$ \\
   \alignmid{\includegraphics[width=0.4\linewidth]{baseline.svg}} && 
   \alignmid{\includegraphics[width=0.3\linewidth]{gain_inc.svg}} 
 \end{tabular}
 \end{center}
 \note[item]{No difference at baseline}
 \note[item]{Impairment of learning}
%
\end{frame}

\begin{frame}{Saturation hypothesis}
%
 \begin{center}
 \begin{tabular}{ccc}
   Error model & \hspace{0.1\linewidth} & Saturation model \\[1cm]
   \alignmid{\includegraphics[width=0.3\linewidth]{err_model.svg}} &&
   \alignmid{\includegraphics[width=0.3\linewidth]{sat_model.svg}}
 \end{tabular}
 \end{center}
 \note[item]{Older explanation: error model}
 \note[item]{Our model: baseline activity \lto saturation \lto less depression possible}
%
\end{frame}

\begin{frame}{Evidence: level of depression}
%
 \aligntop{\includegraphics[width=0.3\linewidth]{pSer880.svg}}
 \hp
 \parbox[t]{0.6\linewidth}{%
 Basal level of GluR2 phosphorylation at serine 880 in AMPA receptor.

 \vp Biochemical signature of PF-Pk LTD.

 \vp Shows that \# depressed synapses in flocculus is larger in KO than WT.
 \note[item]{Predicted by saturation hypothesis}
 }
%
\end{frame}

\begin{frame}{Evidence: saturation by CF stimulation}
%
 Use Channelrhodopsin to stimulate CF \parbox[t]{0.45\linewidth}{%
 \lto increase LTD in PF-Pk synapses \\
 \lto simulate saturation in WT.
 }
 \note[item]{should result in similar behaviour to KO}

 \vp\begin{center}
   \alignmid{\includegraphics[width=0.4\linewidth]{CFstim.svg}}
   \hspace{0.1\linewidth}
   \alignmid{\includegraphics[width=0.4\linewidth]{CFstim_results.svg}}
 \end{center}
%
\end{frame}

\begin{frame}{Evidence: desaturation by gain-decrease pretraining}
%
 %
 \begin{center}
   \includegraphics[width=0.7\linewidth]{desat.svg}

   \vp \includegraphics[width=0.7\linewidth]{gain_dec.svg}
 \end{center}
 %
 \note[item]{gain dec reverses LTD}
 \note[item]{but behaviour from elsewhere}
%
\end{frame}

\begin{frame}{Summary of training results}
%
 \begin{center}
   \alignmid{\includegraphics[width=0.6\linewidth]{data_shifted.svg}}
   \hspace{0.05\linewidth}
   \alignmid{\includegraphics[width=0.3\linewidth]{comparisons.svg}}
 \end{center}
 \note[item]{Restricted to gain inc for comparison}
 \note[item]{Black: WT. Red: KO}
 \note[item]{Solid: no pre. Dashed: with pre}
 \note[item]{Horz and vert comparisons: conceptual}
 \note[item]{Diagonal comparisons: paramter fitting. Depend on size of KO vs.\ pretraining}
 \note[item]{KO hurts w/o, but helps w/}
 \note[item]{pre helps KO but hurts WT}
 \note[item]{top and left most restrictive}
%
\end{frame}

@@

\begin{frame}{Complex synapses}
%
  \aligntop{\includegraphics[width=5cm]{synapse.svg}}\hp
  %
  \note[item]{functional states, not molecules}
  \note[item]{synaptic weight depends on state}
  \note[item]{many states can have same weight}
  %
  \hp\aligntop{\includegraphics[width=2cm]{pot.svg}}
  \hp\aligntop{\includegraphics[width=2cm]{dep.svg}}
  %
  \note[item]{stochastic transitions}
%
\end{frame}

@@

\begin{frame}{Simplifying assumptions}
%
\begin{itemize}
  \item No spatial/temporal correlations in plasticity events.
      %
      \note[item]{allows us to concentrate on synapse, not neuron/network}
      %
%  \item Which synapses eligible for plasticity chosen randomly.
      %
      \note[item]{No filing system}
      %
  \item Potentiating/depressing plasticity events $\sim$ Poisson processes.% with rates $rf\potdep$, where $f\pot+f\dep=1$.
      %
      \note[item]{don't care if STDP...}
%      %
%      \note[item]{$r=$ total rate of plasticity events per synapse, $f\potdep=$ fraction of events that are potentiating/depressing.}
      %
  \item Potentiation and depression are described by Markov processes.% with transition probabilities $\M\potdep$.
%      %
%      \note[item]{matrix elements: transition prob from $i\to j$, given pot/dep}
%  \item Ideal observer: read weights directly.
%      \note[item]{not electrical activity: don't model neurons/network}
%      \note[item]{upper bound on electrical activity readout}
%      %
%  \item Synaptic weights can only take values $\pm1$.
%      %
%      \note[item]{looks like binary synapse from outside. Inside...}
      %
  \citerr{Fusi2005cascade,Fusi2007multistate,Barrett2008discrete}
\end{itemize}
%
\end{frame}

@@

\begin{frame}{Synaptic dynamics}
%
  There are $N$ identical synapses with $M$ internal functional states.
  %
  \begin{center}
    \includegraphics[width=0.7\linewidth]{model.svg}
  \end{center}
  %
  \note[item]{stoch process has steady state.}
  \note[item]{Prior activity puts it in this state. row vec.}
  \note[item]{plasticity events at rate r}
  \note[item]{fraction pot/dep}
  \note[item]{probs changed by Markov matrices, prob $i\to j$}
  \note[item]{Readout: synaptic weight vec when in each state.}
  \vp
% At $t=0$, the memory is created by $\M\potdep$ with probability $f\potdep$.
% \note[item]{for this one, we keep track of pot/dep, look for inc/dec of $\w$}
%
% \vp Forgetting caused by subsequent memories, evolving as
%
  \begin{equation*}
  \begin{aligned}
    \diff{\pr(t)}{t} &= r\pr(t)\frg,
    &\qquad
    \frg &= f\pot\M\pot+f\dep\M\dep-\I,\\&&
    \eq\frg &=0.
  \end{aligned}
  \end{equation*}
%
  \note[item]{Memory at $t=0$, keep track of pot/dep}
  \note[item]{subsequent: average over pot/dep}
% \note[item]{$\frg$ is forgetting matrix, $\I=$identity, don't keep track of pot/dep}
% Eventually, this will settle into the equilibrium distribution:
% %
% %
% \note[item]{In equilibrium prior to memory creation}
%
\end{frame}

\begin{frame}{Modelling VOR learning}
%
\begin{tabular}{>{\bfseries}ll>{\impl}l}
    Mutation: &
    Changes mechanism of LTD  &
    change $\M\dep$.
    \note[item]{lower threshold \lto increase off-diagonal elements.}
 \\[1cm]
    Training: &
    Changes statistics of LTP/LTD  &
    change $r,f\pot,f\dep$.
    \note[item]{Only parameters we have. Don't care about $r$.}
    \note[item]{Same PF+CF input \lto same $r,f\pot,f\dep$ in each case.}
 \\[1cm]
    Learning: &
    Change in VOR gain &
    fn.\ of decrease in $\av{\w}$.
    \note[item]{Only output we have. Don't keep track of synaptic identity.}
    \note[item]{Input to Pk, some linear combination of $\w$'s. }
\end{tabular}
%
\end{frame}

@@

\begin{frame}{Serial synapse}
%
 %
 \begin{center}
   \includegraphics[width=0.5\linewidth]{serial.svg}\\
   \citerr{Leibold2008serial} \\[1cm]
 \note[item]{Still looks binary from outside. Hidden states (not essential).}
 \note[item]{Only see $\Delta w$ at boundary.}
   \aligntop{\includegraphics[width=0.35\linewidth]{serial_learnS.eps}}
   \hspace{2cm}
   \aligntop{\includegraphics[width=0.35\linewidth]{comparisons_serial.svg}}
 \end{center}
 %

@@

%%-------------Slide--------------------------------------------------------
%
%\begin{frame}{Acknowledgements}
%%
% 
%%
%\end{frame}
%
%-------------Slide--------------------------------------------------------
